The Effect of Oil Price on Field Production: Evidence from the Norwegian Continental Shelf

### 2016-03-03 oil_prices_obes.tex
update

The effect of oil prices on producing fields is an important topic in understanding the mechanisms of how total oil supply reacts in response to price. The response of oil production in existing fields is especially important now as many of the major oil-producing areas, like the Norwegian Continental Shelf are mature and an increasing share of investments will be directed at capacity enhancement in producing fields rather than new field developments. How production from these fields will respond to changes in oil prices has major implications for the oil industry, the state finances of oil-producing countries and long-run oil price formation.

@@

An expanding litterature looking at more microeconomic issues of oil supply exists. \citet{kellogg_effect_2014} studies the effect of oil-price uncertainty on drilling and exploration. The author finds that oil exploration firms in Texas do approximately respond as real-option theory would predict when it comes to the timing of drilling.  A model and test using data from North Sea producers on the UK continental shelf by \citet{hurn_geology_1994}, on the other hand, fails to find evidence that the variance in the oil price affects the timing of oil field development.

>>>>

The effect of oil prices on producing fields is an important topic in understanding the mechanisms of how total oil supply reacts in response to price. The response of oil production in existing fields is especially important now as many of the major oil-producing areas, like the Norwegian Continental Shelf are mature and an increasing share of investments will be directed at capacity enhancement in producing fields rather than new field developments. Figure \ref{investment} shows that investments in producing fields makes up a large share of total investments, nearly dwarfing investments in searching, exploration and new build-outs. How production from these fields will respond to changes in oil prices has major implications for the oil industry, the state finances of oil-producing countries and long-run oil price formation.

\begin{figure}
	\includegraphics[width=1\textwidth]{figures/investment.png}
	\caption{The figure shows investments (millions NOK) in the oil and gas extraction industry on the Norwegian Continental Shelf. A large and growing share of investment goes into fields that are already producing as opposed to investments in search, exploration and build-out of new fields.}
	\label{investment}
\end{figure}

@@

An expanding litterature looking at more microeconomic issues of oil supply exists. \citet{kellogg_effect_2014} studies the effect of oil-price uncertainty on drilling and exploration. The author finds that oil exploration firms in Texas approximately respond as real-option theory would predict when it comes to the timing of drilling.  A model and test using data from North Sea producers on the UK continental shelf by \citet{hurn_geology_1994}, on the other hand, fails to find evidence that the variance in the oil price affects the timing of oil field development.
